---
title: "Import and Tibbles"
output: html_document
---
  
  ```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r libraries, echo = TRUE, results='hide', message=FALSE}
library(tidyverse)
```

### Chapter 8 - Data Import with readr

#### Getting Started

##### Examples 

```{r eval=FALSE}
read_csv("a,b,c
1,2,3
4,5,6")
```
```{r eval = FALSE}
 read_csv("The first line of metadata
          The second line of metadata
          x,y,z
          1,2,3", skip = 2)
```
```{r eval = FALSE}
 read_csv("# A comment I want to skip
          x,y,z
1,2,3", comment = "#")
```
```{r eval = FALSE}
read_csv("1,2,3\n4,5,6", col_names = FALSE)
```
```{r eval = FALSE}
read_csv("a,b,c\n1,2,.", na = ".")
```




##### Exercises

1. What function would you use to read a file where fields are separated with “|”?
2. Apart from file, skip, and comment, what other arguments do `read_csv()` and `read_tsv()` have in common?
3. What are the most important arguments to `read_fwf()`?
4. Sometimes strings in a CSV file contain commas. To prevent them from causing problems they need to be surrounded by a quoting character, like " or '. By convention, read_csv() assumes that the quoting character will be ", and if you want to change it you’ll need to use `read_delim()` instead. What arguments do you need to specify to read the following text into a data frame?
        "x,y\n1,'a,b'"
5. Identify what is wrong with each of the following inline CSV files. What happens when you run the code?
        `read_csv("a,b\n1,2,3\n4,5,6")`
        `read_csv("a,b,c\n1,2\n1,2,3,4")`
        `read_csv("a,b\n\"1")`
        `read_csv("a,b\n1,2\na,b")`
        `read_csv("a;b\n1;3")`

#### Parsing a vector

##### Examples 

```{r eval = FALSE}
str(parse_logical(c("TRUE", "FALSE", "NA")))
str(parse_integer(c("1", "2", "3")))
str(parse_date(c("2010-01-01", "1979-10-14"))) 
```

```{r eval = FALSE}
parse_integer(c("1", "231", ".", "456"), na = ".")
x <- parse_integer(c("123", "345", "abc", "123.45"))
x
problems(x)
```

```{r eval = FALSE}
parse_double("1.23")
parse_double("1,23", locale = locale(decimal_mark = ","))
```
```{r eval = FALSE}
parse_number("$100")
parse_number("20%")
parse_number("It cost $123.45")
```

```{r eval = FALSE}
# Used in America
parse_number("$123,456,789") 

# Used in many parts of Europe
parse_number(
      "123.456.789",
      locale = locale(grouping_mark = ".")
)
# Used in Switzerland
parse_number(
      "123'456'789",
      locale = locale(grouping_mark = "'")
)
```
```{r eval = FALSE}
charToRaw("Rladies MCR")
```

```{r eval = FALSE}
x1 <- "El Ni\xf1o was particularly bad this year"
x2 <- "\x82\xb1\x82\xf1\x82\xc9\x82\xbf\x82\xcd"


parse_character(x1, locale = locale(encoding = "Latin1"))
parse_character(x2, locale = locale(encoding = "Shift-JIS")) 
```
```{r eval = FALSE}
fruit <- c("apple", "banana")
parse_factor(c("apple", "banana", "bananana"), levels = fruit)
```
```{r eval = FALSE}
parse_datetime("2010-10-01T2010")
```
```{r eval = FALSE}
# If time is omitted, it will be set to midnight
parse_datetime("20101010")
```
```{r eval = FALSE}
library(hms)
parse_time("01:10 am")
parse_time("20:10:01")
```

```{r eval = FALSE}
parse_date("01/02/15", "%m/%d/%y")
parse_date("01/02/15", "%d/%m/%y")
parse_date("01/02/15", "%y/%m/%d")

```

##### Exercises
1. What are the most important arguments to `locale()`? 
2. What happens if you try and set `decimal_mark` and `grouping_mark` to the same character? What happens to the default value of `grouping_mark` when you set `decimal_mark` to ",“? What happens to the default value of `decimal_mark` when you set the grouping_mark to ".“?
3. I didn’t discuss the `date_format` and `time_format` options to `locale()`. What do they do? Construct an example that shows when they might be useful.
4. If you live outside the US, create a new locale object that encapsulates the settings for the types of files you read most commonly.
5. What’s the difference between `read_csv()` and `read_csv2()`?
6. What are the most common encodings used in Europe? What are the most common encodings used in Asia? Do some goo‐ gling to find out.
7. Generate the correct format string to parse each of the following dates and times:
d1 <- "January 1, 2010"
d2 <- "2015-Mar-07"
d3 <- "06-Jun-2017"
d4 <- c("August 19 (2015)", "July 1 (2015)") d5 <- "12/30/14" # Dec 30, 2014
        t1 <- "1705"
        t2 <- "11:15:10.12 PM"
        
#### Parsing a File
##### Exercise

Strategy 1

```{r eval = FALSE}
challenge <- read_csv(readr_example("challenge.csv"))
```
```{r eval = FALSE}
problems(challenge)
```
```{r eval = FALSE}
 challenge <- read_csv(
      readr_example("challenge.csv"),
      col_types = cols(
        x = col_integer(),
        y = col_character()
      )
)
```
```{r eval = FALSE}
challenge <- read_csv(
      readr_example("challenge.csv"),
      col_types = cols(
        x = col_double(),
        y = col_character()
      )
)

```
```{r eval = FALSE}
tail(challenge)
```

```{r eval = FALSE}
 challenge <- read_csv(
      readr_example("challenge.csv"),
      col_types = cols(
        x = col_double(),
        y = col_date()))
```

```{r eval = FALSE}
tail(challenge)
```

Strategy 2

```{r eval = FALSE}
challenge2 <- read_csv(
                        readr_example("challenge.csv"),
                        guess_max = 1001
                      )
```

```{r eval = FALSE}
challenge2 <- read_csv(readr_example("challenge.csv"),
        col_types = cols(.default = col_character())
)
```

#### Writing to a File 
##### Examples 


```{r eval = FALSE}
challenge 
write_csv(challenge, "challenge-2.csv") 
read_csv("challenge-2.csv")
```

```{r eval = FALSE}
write_rds(challenge, "challenge.rds") 
read_rds("challenge.rds")
```

```{r eval = FALSE}
library(feather)
write_feather(challenge, "challenge.feather")
read_feather("challenge.feather")
```

### Chapter 7 - Tibbles with tibble ... 

#### Creating tibbles
##### Examples 
```{r eval = FALSE}
as_tibble(iris)
```
```{r eval = FALSE}
tibble( x=1:5,
y=1,
z=x^2+y )
```
```{r eval = FALSE}
tb <- tibble(
      `:)` = "smile",
      ` ` = "space",
      `2000` = "number"
)
tb
```

#### Tibbles Versus data.frame 
##### Examples 
Printing
```{r eval = FALSE}
tibble(
a = lubridate::now() + runif(1e3) * 86400, b = lubridate::today() + runif(1e3) * 30, c = 1:1e3,
d = runif(1e3),
e = sample(letters, 1e3, replace = TRUE)
)
```

Subsetting

```{r eval=FALSE, include=TRUE}
df <- tibble(
      x = runif(5),
      y = rnorm(5)
)
df
df$x
    # Extract by position
df[[1]]
```

```{r eval=FALSE, include=TRUE}
library(datasets)
austres <- as.tibble(austres)
```

```{r eval=FALSE, include=TRUE}
austres %>% print(n = 10, width = Inf)
```
```{r eval=FALSE, include=TRUE}
austres %>% View()
```

```{r eval=FALSE, include=TRUE}
df %>% .$x
df %>% .[["x"]]
```

#### Interacting with Older Code 
##### Examples 
```{r eval = FALSE}
class(as.data.frame(tb))
```


##### Exercises

1. How can you tell if an object is a tibble? (Hint: try printing mtcars, which is a regular data frame.)
2. Compare and contrast the following operations on a data.frame and equivalent tibble. What is different? Why might the default data frame behaviors cause you frustration?
       `df <- data.frame(abc = 1, xyz = "a")`
       `df$x`
       `df[, "xyz"]`
       `df[, c("abc", "xyz")]`
3. If you have the name of a variable stored in an object, e.g., var <- "mpg", how can you extract the reference variable from a tib‐ ble?
4. Practice referring to nonsyntactic names in the following data frame by:
a. Extracting the variable called 1.
b. Plotting a scatterplot of 1 versus 2.
c. Creating a new column called 3, which is 2 divided by 1.
d. Renaming the columns to one, two, and three:
          annoying <- tibble(
            `1` = 1:10,
            `2` = `1` * 2 + rnorm(length(`1`))
)
5. What does tibble::enframe() do? When might you use it?
6. What option controls how many additional column names are
printed at the footer of a tibble?
